\chapter{Introduction} 

\section{Context}
The \xten robotics platform is a general purpose robotics kit that is designed to be a low cost, modular, scalable and extensible option for developing custom robotics projects. The system's intended audience and applications range from simple primary school level robots to more complex creations used at university level robotics competitions and research programs. The \xten kit's focus is to improve upon current general robotics kit standards by providing:
\begin{itemize}
\item Support for potentially hundreds of sensors and actuators.
\item A modular design with pluggable add-on peripheral boards.
\item An extensible software and hardware architecture that easily supports new devices.
\item A low overall cost while achieving all the above listed requirements.
\end{itemize}

Although many general-purpose robotics kits exist, many are either quite limited in the range of models that can be built with them, or demand a high price for their expressiveness. Many of the current robotics kits are designed with a limited number of available ports for attaching sensors and actuators. This is evident in the general purpose LEGO Mindstorms Robotics kit and most of the wheeled platforms like the Boe-bot and the Rug Warrior Pro. %~\parencite{6frompaper}. 
The underlying designs of these kits only partially support scaling, or not at all: one is forced to limit the features of one's robotic designs to operate within this limitation. Inadequate peripheral interfaces can curb creativity and innovativeness, resulting in unfulfilled designs, especially for robotics novices.

The \xten robotics platform is a mix of both hardware and software. The hardware consists of two types of modules, a motherboard and daughterboard. These however can be implemented on a single piece of hardware with a single microprocessor or on their own physically separate hardware modules. For physically separate boards, they are connected via a bus that passes data and instructions between a single motherboard and one or more daughterboards. The daughterboards in turn process these instruction and data and subsequently execute them by applying them to sensors and actuators connected to their respective ports.

There are specialised software components present on each hardware module. The motherboard hosts the user generated program code and the system software that processes it. The \hyperref[ssub:middleware]{middleware} is present on both the motherboard and daughterboard, and is responsible for seamless cross-board communication. Finally there is the firmware which is embedded in the daughterboard, this is semi-permanent software that is specialised to interpret and execute instructions passing between the sensors and actuators to and from the motherboard.

\section {Motivation}
The motivating factors for this project came from the need for involving
robotics in education. This lead to the creation of the \texttt{X10ABOT} robotics
platform. This project is aimed at creating a platform that will 
adequately provide a full featured robotics framework for students, all at a
price lower than the current market price for a full featured robotics kit.
This will further assist in the commoditization of robotics kits and subsequently
increase the possibility for activities such as robotics competitions between schools.
Other driving motivations for the project include the creation of a platform that
will provide a flexible starting point for other robotics research projects. The kit
will also be useful enough for hobbyist users who are interested in robotics
and automation, the \texttt{X10ABOT} robotics platform will be an attractive tool that
can be utilized in their projects.


We researched the structure of several robotics research projects and commercial robotics kits and identified a number of design objectives that were useful for a general purpose architecture. We aimed at creating a robotics framework that was able to support more sensors and actuators than most of the popular robotics kits, this framework had to support most of the common off-the-shelf sensors and actuators and yet be easy to use without limiting creativity. 

We reviewed many types of robotics projects and came across a few interesting concepts. They influenced and validated the design decisions made in the \xten project. We will review a few of them and present an overview of their architectures. (in chapter 2?)

\section{Thesis Summary}
Outlined in this thesis are the details of how we progressed from our conceptual ideal product of a low-cost, modular, scalable, extensible robotics kit to transforming these ideals into realistic software and hardware components. These components are modelled and tested using the Arduino embedded development platform. We utilize one or more Arduinos to represent the motherboard daughterboard setup that we envisioned. We also take advantage of the Arduino software framework which is used to develop the system software, \hyperref[ssub:middleware]{middleware} and firmware.

The outcome of the use of these tools and strategies gives us a system that is able to easily control multiple sensors and actuators in a scalable and modular way while allowing for extensibility.

Figure~\textbf{\ref{code:simpl}} describes a typical code setup for a very simple mobile robot using the \xten framework. All the sensors and actuators are placed on one board (the daughterboard) for this simple implementation. This robot has one actuator (a DC motor connected to an external H-bridge), a force sensor and a push-button sensor. The instruction in Figure~\textbf{\ref{code:simpl}} drives the motor at full speed if either the touch sensor records an input or if there is a force on the force sensor above a certain threshold, otherwise the motor will be given the off signal.

 We then further extend the same simple example from Figure~\textbf{\ref{code:simpl}} to show how the \xten framework is able to scale modularly.
 In Figure~\ref{code:simplplus} an extra DC motor and a light sensor is added as a single module. Instructions are then added to activate the new DC motor if there is light intensity above a certain threshold on the sensor. The code example shows that even with the addition of two independent devices, \emph{motor1} and \emph{lightSensor} to the system, it did not affect the existing code but fit seamlessly in the development process.
 

\begin{figure}
		\footnotesize
    {\fontsize{8}{6}\selectfont
		\begin{minted}[bgcolor=bg,baselinestretch=1,frame=lines,framesep=2mm,label={Simple Example}]{c}

/**
* Import the necessary libraries for the motherboard, 
* internal daughterboard and the peripheral bus
**/
#include <Wire.h>  
#include <X10ABOT_MB.h>
#include <X10ABOT_DB.h> //Include the internal daughterboard #0 (SELF)

//Initialise the DC motor on daughterboard #0 (SELF), actuator port #1
Actuator motor1(SELF,1);
//Initialise force sensor on daughterboard #0 (SELF), sensor port #1
Sensor force1(SELF,1);
//Initialise force sensor on daughterboard #0 (SELF), sensor port #2
Sensor pushbutton(SELF,1);

void setup(){}
void loop(){
//Continuously check the sensors for a reading
  if(pushbutton.readDigitalB() || (force1.readAnalog()>100)){
    motor1.setDigital(LO, HI); //Turn motor1 on by sending LO on pin a and HI on pin B
    motor1.pwm_a(100); //operate motor1 at full power (100%) 
  }else{
    motor1.setDigital(LO, LO);//Turn motor1 off by sending LO on pin a and pin b
  }
}	 
	\end{minted}
  }
  \caption{Example of a single controller with a force sensor and a DC motor} \label{code:simpl}
		
\end{figure}


\begin{figure}
		\footnotesize
    {\fontsize{8}{6}\selectfont
		\begin{minted}[bgcolor=bg,baselinestretch=1,frame=lines,framesep=2mm,label={Complex addition to simple example}]{c}
        \end{minted}
        \begin{minted}[bgcolor=bg,baselinestretch=1]{c}
/**
* Import the necessary libraries for the motherboard, 
* internal daughterboard and the peripheral bus
**/
#include <Wire.h>  
#include <X10ABOT_MB.h>
#include <X10ABOT_DB.h> //Include the internal daughterboard #0 (SELF)

//Initialise the DC motor on daughterboard #0 (SELF), actuator port #1
Actuator motor1(SELF,1);
//Initialise force sensor on daughterboard #0 (SELF), sensor port #1
Sensor force1(SELF,1);
//Initialise force sensor on daughterboard #0 (SELF), sensor port #2
Sensor pushbutton(SELF,1);
   
void setup(){}
void loop(){
//Continuously check the sensors for a reading
if(pushbutton.readDigitalB() || (force1.readAnalog()>100)){
    motor1.setDigital(LO, HI); //Turn motor1 on by sending LO on pin a and HI on pin B
    motor1.pwm_a(100); //operate motor1 at full power (100%) 
  }else{
    motor1.setDigital(LO, LO);//Turn motor1 off by sending LO on pin a and pin b
  }
 const byte BOARD2 = 9; //Constant with arbitrary daughterboard address
 //Declare motor on daughterboard #9, actuator port #1
 Actuator motor2(BOARD2,1);
 //Declare force sensor on daughterboard #9, sensor port #1
 Sensor lightSensor(BOARD2,1);
 
//Added extra functionality with a new sensor and a new actuator
//on daughterboard #9 while light is on the sensor, activate motor2
if (lightSensor.readAnalog()>700)
  {
    motor2.setDigital(HI, LO); //Turn motor2 on by sending HI on pin A and LO on pin b
    motor2.pwm_a(50); //operate motor2 at half power (50%)
  }else{
    motor2.setDigital(LO, LO);//Turn motor1 off by sending LO on pin a and pin b
  }
}	 
	\end{minted}
  }
		\caption{Now we add: A push-button sensor to the motherboard (or daughterboard \#0). We also add an extra daughterboard with a light sensor, and a DC Motor} \label{code:simplplus}
    
\end{figure}

\paragraph*{Thesis Statement}
\textit{It is possible to design an architecture for a general purpose robotics platform using the Arduino\texttrademark\ microcontroller and supporting software, open-source libraries and tools. It appears that by using a modular, distributed architecture for both hardware and software, it is possible to create a low cost, scalable system with the ability to support hundreds of sensors and actuators. Furthermore, the same API for hardware control that is used to implement the modular and scalable design can be leveraged to allow user defined extensions to hardware interfaces. Finally, all of this can be implemented with only moderate latency penalties (~250-540$\mu$s per digital read/write) .} 


\section{Structure of Thesis}
Chapter 2 explores the design goals of the \xten platform, how the same principles are being used by other robotics systems and how we apply them for our purpose. 
Chapter 3 describes the design and implementation of the entire platform, the technologies used and how we integrated them to work seamlessly together.
Chapter 4 illustrates the results that demonstrates the effectiveness of the design strategies we use.
Chapter 5 explores the potential areas for future work and concludes the thesis.

